
## Fin-Eval: Assignment 3

Rob Alderman  
[Financial Evaluation and Investment Strategy](https://www.coursera.org/learn/investments/outline?module=7TIU9)  
[Assignment 3](https://www.coursera.org/learn/investments/lecture/meglV/assignment-3-analyzing-identifying-three-mystery-securities)  
Sept 2015

<br />
### Data source

[ThreeMysterySecurities.xlsx](ThreeMysterySecurities.xlsx)
from [https://www.coursera.org/learn/investments/supplement/dgfBm/module-2-spreadsheets-and-other-resources](https://www.coursera.org/learn/investments/supplement/dgfBm/module-2-spreadsheets-and-other-resources)  

Source for U.S. Stock Returns and U.S. Treasury Bill Rates: [Kenneth R. French Data Library](http://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html)  
For defintions of Fama-French Factors (Mkt_RF, SMB, & HML), see [here](http://mba.tuck.dartmouth.edu/pages/faculty/ken.french/Data_Library/f-f_factors.html)  

```{r}

# Need to set JAVA_HOME and PATH, for read.xlsx
Sys.setenv(JAVA_HOME="C:\\fox\\java\\jre")
p <- Sys.getenv("PATH")
p1 <- paste("C:\\fox\\java\\jre\\bin\\j9vm;", p, sep="")
Sys.setenv(PATH=p1)

library(xlsx)
data <- read.xlsx("ThreeMysterySecurities.xlsx",
                  sheetIndex=1,
                  rowIndex=4:244,
                  colIndex=c(1:5,7:9))
str(data)
head(data)

data$Security.1_RF <- data$Security.1 - data$RF
data$Security.2_RF <- data$Security.2 - data$RF
data$Security.3_RF <- data$Security.3 - data$RF

```

<br />
### Question 1

Fill in the table for *Mystery Security #1*.  
What's the identity of the mystery security?  What are its characteristics?

```{r}

capm.sec1 <- lm(Security.1_RF ~ Mkt_RF, data=data)
summary(capm.sec1)

fac3.sec1 <- lm(Security.1_RF ~ Mkt_RF + SMB + HML, data=data)
summary(fac3.sec1)

library(ggplot2)
qplot(y=Security.1_RF, x=Mkt_RF, data=data) +
    ggtitle("CAPM - Security 1") + 
    geom_hline(y=0,colour="lightgray") +
    geom_vline(x=0,colour="lightgray") +
    geom_abline(intercept = coef(capm.sec1)["(Intercept)"], 
                slope = coef(capm.sec1)["Mkt_RF"], 
                colour="blue")

```


| Security 1                |  Values                                   |
| ---------------------     | ----------------------------------------- |
| Avg return                | `r mean(data$Security.1)`                 |
| Avg excess return         | `r mean(data$Security.1_RF)`              |
| CAPM alpha                | `r coef(capm.sec1)[["(Intercept)"]]`      |
| CAPM beta                 | `r coef(capm.sec1)[["Mkt_RF"]]`           |
| CAPM R<sup>2</sup>        | `r summary(capm.sec1)$r.squared`          |
| 3-factor alpha            | `r coef(fac3.sec1)[["(Intercept)"]]`      |
| 3-factor market beta      | `r coef(fac3.sec1)[["Mkt_RF"]]`           |
| 3-factor size beta        | `r coef(fac3.sec1)[["SMB"]]`              |
| 3-factor value beta       | `r coef(fac3.sec1)[["HML"]]`              |
| 3-factor R<sup>2</sup>    | `r summary(fac3.sec1)$r.squared`          |


Security 1 has beta > 1.  It amplifies market volatility. alpha < 0, so it underperforms w/r/t market volatility.
Its average return is less than that of the market. 
About 63% of volatility is explained by the market.  The 3-factor model didn't improve R<sup>2</sup>.
The coefficients on size and value are small and the market coefficient hardly changed from the CAPM model.
So this security is a mid-cap stock that's very sensitive to market conditions.

Disaster.  Low return combined with high beta.  Underperforms w/r/t volatility (alpha).  It also
underperforms the market in terms of return.  Looks like a mutual fund (mid-high R<sup>2</sup>) focused
on small-cap staocks (slight correlation with size and value).



<br />
### Question 2

Fill in the table for *Mystery Security #2*.  
What's the identity of the mystery security?  What are its characteristics?

```{r}

capm.sec2 <- lm(Security.2_RF ~ Mkt_RF, data=data)
summary(capm.sec2)

fac3.sec2 <- lm(Security.2_RF ~ Mkt_RF + SMB + HML, data=data)
summary(fac3.sec2)

library(ggplot2)
qplot(y=Security.2_RF, x=Mkt_RF, data=data) +
    ggtitle("CAPM - Security 2") + 
    geom_hline(y=0,colour="lightgray") +
    geom_vline(x=0,colour="lightgray") +
    geom_abline(intercept = coef(capm.sec2)["(Intercept)"], 
                slope = coef(capm.sec2)["Mkt_RF"], 
                colour="blue")
```


| Security 2                |  Values                                   |
| ---------------------     | ----------------------------------------- |
| Avg return                | `r mean(data$Security.2)`                 |
| Avg excess return         | `r mean(data$Security.2_RF)`              |
| CAPM alpha                | `r coef(capm.sec2)[["(Intercept)"]]`      |
| CAPM beta                 | `r coef(capm.sec2)[["Mkt_RF"]]`           |
| CAPM R<sup>2</sup>        | `r summary(capm.sec2)$r.squared`          |
| 3-factor alpha            | `r coef(fac3.sec2)[["(Intercept)"]]`      |
| 3-factor market beta      | `r coef(fac3.sec2)[["Mkt_RF"]]`           |
| 3-factor size beta        | `r coef(fac3.sec2)[["SMB"]]`              |
| 3-factor value beta       | `r coef(fac3.sec2)[["HML"]]`              |
| 3-factor R<sup>2</sup>    | `r summary(fac3.sec2)$r.squared`          |


Security 2 has excellent returns along with politive alpha and a small beta. 
So it outperforms w/r/t market volatility (alpha > 0).
Its a defensive stock - it correlates with market volatility but in a muted fashion.  
The R<sup>2</sup> is woefully small, so most of the security's volatility is idiosyncratic.
The 3 factor model showed a negative correlation with size and positive correlation with value.
So this appears to be a Large-cap Value stock, necessity good (low beta)

<br />
### Question 3

Fill in the table for *Mystery Security #3*.  
What's the identity of the mystery security?  What are its characteristics?

```{r}

capm.sec3 <- lm(Security.3_RF ~ Mkt_RF, data=data)
summary(capm.sec3)

fac3.sec3 <- lm(Security.3_RF ~ Mkt_RF + SMB + HML, data=data)
summary(fac3.sec3)

library(ggplot2)
qplot(y=Security.3_RF, x=Mkt_RF, data=data) +
    ggtitle("CAPM - Security 3") + 
    geom_hline(y=0,colour="lightgray") +
    geom_vline(x=0,colour="lightgray") +
    geom_abline(intercept = coef(capm.sec3)["(Intercept)"], 
                slope = coef(capm.sec3)["Mkt_RF"], 
                colour="blue")
```


| Security 3                |  Values                                   |
| ---------------------     | ----------------------------------------- |
| Avg return                | `r mean(data$Security.3)`                 |
| Avg excess return         | `r mean(data$Security.3_RF)`              |
| CAPM alpha                | `r coef(capm.sec3)[["(Intercept)"]]`      |
| CAPM beta                 | `r coef(capm.sec3)[["Mkt_RF"]]`           |
| CAPM R<sup>2</sup>        | `r summary(capm.sec3)$r.squared`          |
| 3-factor alpha            | `r coef(fac3.sec3)[["(Intercept)"]]`      |
| 3-factor market beta      | `r coef(fac3.sec3)[["Mkt_RF"]]`           |
| 3-factor size beta        | `r coef(fac3.sec3)[["SMB"]]`              |
| 3-factor value beta       | `r coef(fac3.sec3)[["HML"]]`              |
| 3-factor R<sup>2</sup>    | `r summary(fac3.sec3)$r.squared`          |


Performing at its CAPM benchmark (alpha = 0).  
Highly correlated with growth stocks.
High R<sup>2</sup> in 3-factor model.
Looks like an actively managed mutual fund of tech stocks.


<br />
### Question 4
For each mystery security, does it have a small or large amount of idiosyncratic risk?

Security 1:   
CAPM R<sup>2</sup> = 63%   
3-factor R<sup>2</sup> = 63%   

About 37% of security 1's volatility is due to idiosyncratic (asset-specific) risk. 

Security 2:  
CAPM R<sup>2</sup> = 8.3%   
3-factor R<sup>2</sup> = 15%   

About 85-90% of security 2's volatility is due to idiosyncratic (asset-specific) risk. 

Security 3:  
CAPM R<sup>2</sup> = 67%   
3-factor R<sup>2</sup> = 82%   

About 20-30% of security 3's volatility is due to idiosyncratic (asset-specific) risk. 



<br />
### Question 5
For each mystery security, what portfolio combination of US Stock Market and Treasury Bills will
yield the same CAPM BETA? 

CAPM BETA is the correlation between the asset's excess returns and the market's excess returns.
What combination of Mkt + RF yields the same correlation?

$$
\begin{align*}
[(w_1 \cdot Mkt + w_2 \cdot RF) - RF] &= \beta [ Mkt - RF ]
\\\\
(w_1 \cdot Mkt + (1 - w_1) \cdot RF)  &= RF + \beta [ Mkt - RF ]
\\\\
w_1 \cdot Mkt + RF - w_1 \cdot RF  &= RF + \beta \cdot Mkt - \beta \cdot RF
\\\\
w_1 \cdot Mkt - w_1 \cdot RF  &= \beta \cdot Mkt - \beta \cdot RF
\end{align*}
$$

So in a portfolio combining the US Stock Market and Treasury Bills, the weight of US Stock Market
allocation (\(w_1\)) corresponds to the CAPM beta for the portfolio.
